---
title: "Classification of Diabetes Pima Indian patients using statistical and machine learning modelling techniques"
output: 
    html_document:
      toc: true
      toc_depth: 2
      number_sections: true
      fig_width: 10
      fig_height: 6
      fig_caption: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if(!require(pacman)) install.packages("pacman", repos = "http://cran.us.r-project.org")

pacman::p_load(tidyverse,dplyr, caret,  stringr, stringi,
              lubridate, skimr, gtsummary, janitor, flextable, # tables design
              scales, knitr, ggthemes, RKaggle,
              corrplot, dlookr, patchwork, tinytex,  MLeval, kknn, nnet, kableExtra,
              plotly, corrr, mice)

#import the data using RKaggle - directly from Kaggle website
data <- get_dataset("uciml/pima-indians-diabetes-database")

```

Author: Shardha Ramlal\
\
Date: `r Sys.Date()`

# Introduction

According to the World Health Organization, type 2 diabetes or diabetes
mellitus occurs when body is unable to use the insulin it produces,
which can lead to high blood sugar if left untreated. 1 The Pima Indian
Population in Arizona has a high incidence of Diabetes, and as such has
been studied extensively for this condition. 2,3 Factors that affect
diabetes include genetic predisposition, not exercising enough and
obesity.1 The research question of the study was to compare
classification models for diabetes to determine the best model.

# Methodology

The dataset used in the study was the Pima Indians Diabetes Database4
was originally from the National Institute of Diabetes and Kidney
Disease, from the study conducted by Smith et al in 1988 2. The dataset
was accessed via the url:
<https://www.kaggle.com/datasets/uciml/pima-indians-diabetes-database/data>
Inclusion criteria was defined as: females with an age greater than or
equal to 21 years at the time of index examination. One examination was
selected per subject – either diabetes diagnosed within five years or,
an oral glucose tolerance test five or more years later failed to reveal
diabetes mellitus.\
Exclusion criteria was defined as if diabetes occurred within one year
of an examination.

Variables in the dataset under study were

-   Pregnancies – Number of times pregnant

-   Glucose – Plasma glucose concentration a 2 hours in an oral glucose
    tolerance test

-   BloodPressure - Diastolic blood pressure (mm Hg)

-   SkinThickness - Triceps skin fold thickness (mm)

-   Insulin - 2-Hour serum insulin (mu U/ml)

-   BMI - Body mass index (weight in kg/(height in m)\^2

-   DiabetesPedigreeFunction - Diabetes pedigree function which provides
    a measure of expected genetric influence of eventual diabetes risk

-   Age - Age (years)

-   Outcome - Class variable (0 or 1) 268 of 768 are 1 (Diabetic), the
    others are 0 (Not Diabetic)

After examination of the data, some variables were missing – these were
coded as zero’s in the dataset and included the variables: Glucose,
Blood Pressure, Skin Thickness, Insulin, BMI. Values of zero were
handled using imputation, similar to study done by Nassiwa and Zeng in
2024 5 however this was done in R using the mice6 package with
predictive mean matching . Ten multiple imputations were used of which
the first was chosen.

The dataset was then split into a 70 % training set and a 30 % test set.
Predictors were scaled, centred and zero variance predictors were
removed. To compensate for the class imbalance in the dataset,
stratified 10 fold cross validation was used Supervised methods of
binary classification consisting of both statistical and machine
learning models were used. These included:

-   Logistic Regression

-   Elastic Net Regression

-   Linear Discriminant Analysis

-   Decision Tree

-   Random Forest

-   Support Vector Machine (Linear Kernel)

-   Support Vector Machine (Radial Kernel)

-   Neural Network

-   K Nearest Neighbours

Best models were then considered using  Receiver Operating Characteristic curve (ROC), Matthew's Correlation Coefficient (normalised version was considered to ensure the metrics were on the same scale), and Accuracy.

# Results

## Data Description and Visualisation

### Missing data in dataset

```{r, echo = FALSE , warning=FALSE, results='asis'}

#change outcome to factor

data$Outcome <- as.factor(data$Outcome)

data$Outcome <- recode(data$Outcome,
                       "1" = "diabetic",
                       "0" = "not_diabetic")

#explore the dataset

# graphs
# quality <- skim(data)
# quality 

###############################################################################

# str(data)

# Some 0's in Glucose, Blood Pressure, Skin Thickness, Insulin, BMI

#replace with NA's to work with

data$Glucose[data$Glucose == 0] <- NA
data$SkinThickness[data$SkinThickness == 0] <- NA
data$BloodPressure[data$BloodPressure == 0] <- NA
data$Insulin[data$Insulin == 0] <- NA
data$BMI[data$BMI == 0] <- NA

# Isolate these for further observation

# kable(colSums(is.na(data[, c("Glucose", "SkinThickness","BloodPressure","Insulin","BMI")])),
#       col.names = "Missing Rows")

#run and extracted

# diagnose_web_report(data, output_format = "html")
# 
# 
# data |>
#   eda_web_report(target = "Outcome", subtitle = "diabetes",
#                  output_dir = getwd(),
#                  output_file = "EDA_Diabetes.html", theme = "blue",
#                  browse=TRUE)
# 
# 
# data |>
#   diagnose_web_report(target = "Outcome", subtitle = "diabetes",
#                  output_dir = getwd(),
#                  output_file = "Diagnose_Diabetes.html", theme = "blue",
#                  browse=TRUE)

missing_pareto <- plot_na_pareto(data, only_na=TRUE, col="blue") +  
  theme(text = element_text(size = 15)) 

ggplotly(missing_pareto)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}

imp_data <- mice(data, m=10, seed = 1918)
# imp_data - pmm used for Glucose, BloodPressure, SkinThickness, Insulin, BMI

# Extract the first imputed dataset
data <- complete(imp_data, 1)

```

### Summary Statistics after Imputation

```{r, echo=FALSE, warning=FALSE, message=FALSE}

data|> tbl_summary(include=everything(),
                    by=Outcome,
                    type = all_continuous() ~ "continuous2" , 
                    statistic = list(all_continuous2() ~ c(
                      "{mean} ({sd})",
                      "{median} ({p25} {p75})",
                      "{min}, {max}"),
                      missing = "no") ,
                    label = list(
                      DiabetesPedigreeFunction = "Diabetes Pedigree Function",
                      BloodPressure = "Blood Pressure" ))|>
        bold_labels() |>  modify_header(
    stat_1 ~ "**Not Diabetic**", stat_2 ~ "**Diabetic**") |>
        as_flex_table() |> autofit() |> 
  font(fontname = "Times New Roman", part = "all")


```

### Overview of Diabetic Patients in the Full Dataset

```{r, echo = FALSE , warning=FALSE, messages=FALSE}

colors <- c('rgb(211,94,96)', 'rgb(128,133,133)')
labels = c('Not Diabetic','Diabetic')

data |> group_by(Outcome) |> summarise(n=n()) |> 
  mutate(perc = n / sum(n)) |> 
  plot_ly(labels = ~labels, values = ~ perc, type = 'pie',
          marker=list(colors=colors))|>
  layout(xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         legend = list(orientation = 'h', x = 0.5, y = -0.1, xanchor = 'center'))



```

### Correlation of Predictors

```{r, echo = FALSE , warning=FALSE, message=FALSE}

corr_matrix <- data |>
  correlate() |>
  as_matrix(diagonal = 1)

plot_ly(
  z = corr_matrix,
  x = colnames(corr_matrix),
  y = rownames(corr_matrix),
  type = "heatmap"
)

```

## Model Building

```{r echo = FALSE, results='hide', fig.show='hide' , message=FALSE, warning=FALSE, include=FALSE}

# #Split dataset into training and test dataset
 set.seed(1918)

# #check for zero variance predictors
nzv <- nearZeroVar(data, saveMetrics= TRUE)


data_subset <- data[,-1] #remove the Outcome (no duplicates)

index <- createDataPartition(data$Outcome,p=0.7,list=FALSE)

train_raw <- data[index,]
test_raw <- data[-index,]

#Scale predictors by subtracting mean and dividing by standard deviation 

preProcValues <- preProcess(train_raw, method = c("center", "scale"))

train <- predict(preProcValues, train_raw)
test <- predict(preProcValues, test_raw)

#remove the unused datasets
rm(test_raw, train_raw, preProcValues, index, nzv)

#Deal with imbalance dataset by stratified cross validation
cvIndex <- createFolds(train$Outcome, 10, returnTrain = TRUE)

fitControl <- trainControl(index=cvIndex,
                           method = "cv", 
                           number = 10,
                           classProbs = TRUE,
                           savePredictions = TRUE,
                           summaryFunction = defaultSummary )


```

```{r echo = FALSE, results='hide', fig.show='hide' , message=FALSE, warning=FALSE, include=FALSE, label=models}

set.seed(1918)
model.lr<- train(Outcome ~. , data = train,
                        method = "glm",
                        trControl = fitControl)

model.lr

summary(model.lr)

model.lr.perf<-confusionMatrix.train(model.lr,norm="none")
model.lr.perf


res.lr <- evalm(model.lr,rlinethick = 1,
                        title="Logistic Regression ROC for data",
                      gname="LR")
res.lr$roc

res.lr$stdres

varImp(model.lr)

model.lr$finalModel

# Elastic Net Regression

set.seed(1918)

model.en <- train(Outcome ~. , data = train,
                        method = "glmnet", family="binomial", 
                        trControl = fitControl)

model.en

summary(model.en)

model.en.perf<-confusionMatrix.train(model.en,norm="none")
model.en.perf


res.en <- evalm(model.en,rlinethick = 1,
                        title="Elastic Net ROC for data",
                      gname="Elastic Net")
res.en$roc

res.en$stdres

varImp(model.en)

# model.en$finalModel

# Linear Discriminant Analysis


set.seed(1918)
model.lda <- train(Outcome ~. , data = train,
                        method = "lda",
                        trControl = fitControl)

model.lda

summary(model.lda)

model.lda.perf<-confusionMatrix.train(model.lda ,norm="none")
model.lda.perf


res.lda <- evalm(model.lr,rlinethick = 1,
                        title="Linear Discriminant Analysis ROC for data",
                      gname="Linear Discriminant")
res.lda$roc

res.lda$stdres

# varImp(model.lda)

model.lda$finalModel

# Decision Tree

set.seed(1918)
model.dt <-  train(Outcome ~. , data = train,
                        method = "rpart",
                        trControl = fitControl)

model.dt

summary(model.dt)

varImp(model.dt)

model.dt.perf<- confusionMatrix.train(model.dt,norm="none")


model.dt.perf

res.dt <- evalm(model.dt,rlinethick = 1,
                     title="Decision Tree ROC for data",
                     gname="Decision Tree")
res.dt$roc
res.dt$stdres

# Random Forest


set.seed(1918)

model.rf <- train(Outcome ~. , data = train,
                        method = "rf",
                        trControl = fitControl)

model.rf

summary(model.rf)

model.rf.perf<-confusionMatrix.train(model.rf,norm="none")
model.rf.perf


res.rf <- evalm(model.rf,rlinethick = 1,
                        title="Random Forest ROC for data",
                      gname="Random Forest")
res.rf$roc

res.rf$stdres

varImp(model.rf)

model.rf$finalModel

# Support Vector Machine with Linear Kernel

set.seed(1918)

model.svm.linear  <- train(Outcome ~. , data = train,
                        method = "svmLinear",
                        trControl = fitControl)

model.svm.linear 

summary(model.svm.linear)


model.svm.linear.perf<-confusionMatrix.train(model.svm.linear,norm="none")
model.svm.linear.perf


res.svm.linear <- evalm(model.svm.linear,rlinethick = 1,
                        title="Support Vector Machine with Linear Kernel ROC for data",
                      gname="SVM Linear")
res.svm.linear$roc

res.svm.linear$stdres

# varImp(model.svm.linear)
# 
# ?VarImp

model.svm.linear$finalModel


# Support Vector Machine with Radial Basis Function Kernel

set.seed(1918)

model.svm.radial  <- train(Outcome ~. , data = train,
                        method = "svmRadial",
                        trControl = fitControl)

model.svm.radial 

summary(model.svm.radial)

model.svm.radial.perf<-confusionMatrix.train(model.svm.radial,norm="none")
model.svm.radial.perf


res.svm.radial <- evalm(model.svm.radial,rlinethick = 1,
                        title="Support Vector Machine with Radial Kernel ROC for data",
                      gname="SVM Radial")
res.svm.radial$roc

res.svm.radial$stdres

# varImp(model.svm.radial)
# 
# ?VarImp

model.svm.radial$finalModel

# Neural Network

set.seed(1918)

model.nn <- train(Outcome ~. , data = train,
                        method = "nnet",
                        trControl = fitControl)

model.nn

summary(model.nn)

model.nn.perf<-confusionMatrix.train(model.nn,norm="none")
model.nn.perf


res.nn <- evalm(model.nn,rlinethick = 1,
                        title="Neural Network ROC for data",
                      gname="Neural Network")
res.nn$roc

res.nn$stdres

varImp(model.nn)


model.nn$finalModel


# K Nearest Neighbours

set.seed(1918)

model.knn <- train(Outcome ~. , data = train, method = "knn",
                        trControl = fitControl)

model.knn

summary(model.knn)

model.knn.perf<-confusionMatrix.train(model.knn,norm="none")
model.knn.perf


res.knn <- evalm(model.knn,rlinethick = 1,
                        title="K Nearest Neighbours ROC for data",
                      gname="K Nearest Neighbours")
res.knn$roc

res.knn$stdres

# varImp(model.knn)


model.knn$finalModel

```

## Comparison of Accuracy of Training Models

```{r echo = FALSE, warning=FALSE, label=results }

metrics <- resamples(
  list(Logistic_Regression = model.lr,
       Elastic_Net = model.en,
       Linear_Discriminant_Analysis = model.lda,
       Decision_Tree=model.dt,
       Random_Forest = model.rf, 
       SVM_linear = model.svm.linear, 
       SVM_radial = model.svm.radial,
       Neural_Network = model.nn,
       K_Nearest_Neighbours=model.knn
  )
)

res_summary <- summary(metrics)

acc <- as.data.frame(res_summary$statistics$Accuracy) 
rownames(acc) <- c("Logistic Regression","Elastic Net",
                           "Linear Discriminant Analysis",
                           "Decision Tree",
                           "Random Forest",
                           "Support Vector Machine (Linear Kernel)",
                           "Support Vector Machine (Radial Kernel)",
                           "Neural Network",
                           "K Nearest Neighbours")


acc |>
  kable(digits = 3, caption = "Model Comparison Metric: Accuracy (CV Folds)") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) |>
  column_spec(1, bold = TRUE) 


# dotplot(metrics) 



```

## Comparison of ROC of Training Models

```{r echo = FALSE, warning=FALSE, message=FALSE ,label=roc_results}
# ROC CURVE ANALYSIS 

models <- list("Logistic Regression" = model.lr,
       "Elastic Net" = model.en,
       "Linear Discriminant Analysis" = model.lda,
       "Decision Tree"=model.dt,
       "Random Forest" = model.rf, 
       "Support Vector Machine (Linear Kernel)" = model.svm.linear, 
       "Support Vector Machine (Radial Kernel)" = model.svm.radial,
       "Neural Network" = model.nn,
       "K Nearest Neighbours"=model.knn
)

# models <- models[order(names(models))]

evalm(models,gnames = names(models),
  rlinethick = 0.8, fsize = 8, plots = "r", silent=TRUE, showplots = FALSE)$roc|> ggplotly()

```

## Performance Metrics of Training Models Summary

```{r echo = FALSE, warning=FALSE ,label=tab_results}

metrics.2 <- cbind(LR=res.lr$stdres,
                         EN=res.en$stdres,
                         LDA=res.lda$stdres,
                         DT=res.dt$stdres,
                         RF=res.rf$stdres,
                         SVM.linear=res.svm.linear$stdres,
                         SVM.radial=res.svm.radial$stdres,
                         NN=res.nn$stdres,
                         KNN=res.knn$stdres
)


metrics.2.test <- as.data.frame(
  cbind(LR=metrics.2[[1]],
        EN=metrics.2[[2]],
        LDA=metrics.2[[3]],
        DT=metrics.2[[4]],
        RF=metrics.2[[5]],
        SVM.linear=metrics.2[[6]],
        SVM.radial=metrics.2[[7]],
        NN=metrics.2[[8]],
        KNN=metrics.2[[9]]))

metrics.2.test   <- tibble::rownames_to_column(metrics.2.test, var = "Metrics")

#remove CI's for clean output
metrics.2.test <- metrics.2.test |> select(Metrics, LR.Score, EN.Score,LDA.Score,DT.Score,RF.Score,SVM.linear.Score,SVM.radial.Score, NN.Score, KNN.Score)

names(metrics.2.test) <- c("Metric","Logistic Regression",
                           "Elastic Net",
                           "Linear Discriminant Analysis",
                           "Decision Tree",
                           "Random Forest",
                           "Support Vector Machine (Linear Kernel)",
                           "Support Vector Machine (Radial Kernel)",
                           "Neural Network",
                           "K Nearest Neighbours")
metrics.2.test |>
  kable(digits = 4, caption = "Model Comparison Metrics (0.5 cutoff)") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) |>
  column_spec(1, bold = TRUE) 





```


# Performance Metrics of Test Dataset

## Exploring the Accuracy of the Test Dataset

```{r warning=FALSE, echo =FALSE, message=FALSE}
#accuracy table 
set.seed(1918)

test_acc <- cbind(
  confusionMatrix(predict(model.lr ,test), test$Outcome)$overall[1],
  confusionMatrix(predict(model.en ,test), test$Outcome)$overall[1],
  confusionMatrix(predict(model.lda ,test), test$Outcome)$overall[1],
  confusionMatrix(predict(model.svm.linear ,test), test$Outcome)$overall[1],
  confusionMatrix(predict(model.nn ,test), test$Outcome)$overall[1]
)

colnames(test_acc) <- c("Logistic Regression",
                           "Elastic Net",
                           "Linear Discriminant Analysis",
                           "Support Vector Machine (Linear Kernel)",
                           "Neural Network")

test_acc |>
  kable(digits = 3, caption = "Model Comparison Metric: Accuracy (Test dataset)") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) |>
  column_spec(1, bold = TRUE)
```


## Comparison of ROC of Test Models

```{r warning=FALSE, echo = FALSE, message=FALSE}
# predicting the test data

set.seed(1918)

test_prob_lr <- predict(model.lr ,test,  type="prob")
test_prob_lr$obs <- test$Outcome
test_prob_lr$Group <- "Logistic Regression"


set.seed(1918)

test_prob_en <- predict(model.en ,test, type="prob")
test_prob_en$obs <- test$Outcome
test_prob_en$Group <- "Elastic Net"

set.seed(1918)

test_prob_lda <- predict(model.lda ,test,  type="prob")
test_prob_lda$obs <- test$Outcome
test_prob_lda$Group <- "Linear Discriminant Analysis"

set.seed(1918)

test_prob_svm.linear <- predict(model.svm.linear ,test,  type="prob")
test_prob_svm.linear$obs <- test$Outcome
test_prob_svm.linear$Group <- "Support Vector Machine (Linear Kernel)"

set.seed(1918)

test_prob_nn <- predict(model.nn ,test, type="prob")
test_prob_nn$obs <- test$Outcome
test_prob_nn$Group <- "Neural Network"

test_comparison <- rbind(test_prob_lr, test_prob_en, test_prob_lda,
                         test_prob_svm.linear, test_prob_nn)


evalm(test_comparison, rlinethick = 0.8, fsize = 8, plots = "r", silent=TRUE, showplots = FALSE)$roc |> ggplotly()



```

## Performance Metrics of Test Models Summary

```{r warning=FALSE, echo = FALSE, message=FALSE}

res_test_comparison <- evalm(test_comparison, showplots = FALSE)

final_metrics_test <- do.call(cbind, res_test_comparison$stdres)


numeric_cols <- sapply(final_metrics_test, is.numeric)

final_metrics_test[numeric_cols] <- round(final_metrics_test[numeric_cols], 3)


final_metrics  <- tibble::rownames_to_column(final_metrics_test, var = "Metrics") |> 
  select(Metrics, `Logistic Regression.Score`,
                                         `Elastic Net.Score`,
                                         `Linear Discriminant Analysis.Score`,
                                         `Support Vector Machine (Linear Kernel).Score`,
                                         `Neural Network.Score`)
                                         
names(final_metrics) <- c("Metric","Logistic Regression",
                           "Elastic Net",
                           "Linear Discriminant Analysis",
                           "Support Vector Machine (Linear Kernel)",
                           "Neural Network")
final_metrics |>
  kable(digits = 3, caption = "Final Model Comparison Metrics") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) |>
  column_spec(1, bold = TRUE)

```

# Discusssion

We used the Receiver Operating Characteristic (ROC) due to the imbalanced nature of the dataset as the best metric for comparison. The best models appeared to be logistic regression, elastic net regression, linear discriminant analysis and support vector machine with a linear kernel. These models had an ROC of 0.84.  

The second metric considered, was the Matthew’s Correlation Coefficient (MCC). We considered the normMCC, which was the MCC scaled within the [0,1]7, yielding 1 as the best value and 0 as the worst value. High MCC values always correspond to high rates in sensitivity, specificity, precision and negative predictive value 7. 

The highest MCC was from the Neural Network model at 0.516 (normMCC = 0.758), followed by the elastic net regression model at 0.483 (normMCC = 0.7415) and the support vector machine with linear kernel model at 0.471 (normMCC = 0.7355) Both logistic regression and linear discriminant analysis models had values of 0.469 (normMCC = 0.7345)

Lastly looking at models according to accuracy metric are the neural network (78.3 %), elastic net regression (77.1 %), support vector machine with linear kernel (76.8%) , linear discriminant analysis (77.1%) and lastly logistic regression (76.6%).

The models that underwent further exploration with the test set were the logistic regression, elastic net regression, linear discriminant analysis, support vector machine with linear kernel and the neural network. 

After further exploration using the hold out set or test set, the highest ROC was still 0.84, however only Logistic Regression and Linear Discriminant analysis kept their performance consistent with the training set. The rest of models reduced performance at 0.83, except for the neural network which remained the same. 

The second metric, normMCC, was highest in the test set for the support vector machine with linear kernel model at 0.7445, higher than its training set value. The elastic net regression model was 0.739, lower than its training data value. The neural network was next at 0.7365, lower than training data performance, followed by logistic regression at 0.735, which was higher than its training data performance. Lastly was the linear discriminant analysis at 0.729 which was lower than its training data performance.  

Lastly looking at models performance according to the accuracy metric – the support vector machine with linear kernel had a performance of 77.8% accuracy, Elastic net had the second highest accuracy, at 77.4%, logistic regression and neural network had 77% accuracy and linear discriminant analysis has 76.5% accuracy. 


# Conclusion

The best model seem to be logistic regression – maintains higher accuracy, ROC consistency, higher normMCC than training data, also this was the simplest model, implying performance will be consistent on new datasets. 

A more complex model, that gave good results was the support vector machine with linear using the normMCC value and accuracy, even better than the training dataset, which implies it will give good performance using unseen data. ROC was slightly lower than the training dataset. 


# References
1. 	Diabetes [Internet]. [cited 2026 Feb 15]. Available from: https://www.who.int/news-room/fact-sheets/detail/diabetes
2. 	Smith JW, Everhart JE, Dickson WC, Knowler WC, Johannes RS. Using the ADAP Learning Algorithm to Forecast the Onset of Diabetes Mellitus. Proc Annu Symp Comput Appl Med Care. 1988 Nov 9;261–5. 
3. 	Pearson ER. Dissecting the Etiology of Type 2 Diabetes in the Pima Indian Population. Diabetes. 2015 Nov 18;64(12):3993–5. 
4. 	Pima Indians Diabetes Database [Internet]. [cited 2025 Nov 16]. Available from: https://www.kaggle.com/datasets/uciml/pima-indians-diabetes-database/data
5. 	Nassiwa F, Zeng J. Evaluating Traditional Machine Learning Models for Predicting Diabetes Onset Using the Pima Indians Dataset. Annals of Medical and Health Sciences Research [Internet]. 2024 [cited 2026 Feb 9]; Available from: https://www.amhsr.org/abstract/evaluating-traditional-machine-learning-models-for-predicting-diabetes-onset-using-the-pima-indians-dataset-12913.html
6. 	Tüzen MF. Handling Missing Data in R: A Comprehensive Guide | R-bloggers [Internet]. 2025 [cited 2026 Feb 9]. Available from: https://www.r-bloggers.com/2025/08/handling-missing-data-in-r-a-comprehensive-guide/
7. 	Chicco D, Jurman G. The Matthews correlation coefficient (MCC) should replace the ROC AUC as the standard metric for assessing binary classification. BioData Mining. 2023 Feb 17;16(1):4. 
8. 	The caret Package [Internet]. [cited 2025 Nov 18]. Available from: https://topepo.github.io/caret/index.html
9. 	Presentation-Ready Data Summary and Analytic Result Tables [Internet]. [cited 2026 Jan 4]. Available from: https://www.danieldsjoberg.com/gtsummary/
10. 	Scatter Plot Matrices - R Base Graphs - Easy Guides - Wiki - STHDA [Internet]. [cited 2026 Jan 4]. Available from: https://www.sthda.com/english/wiki/scatter-plot-matrices-r-base-graphs
11. 	Li D. Chapter 9 Pie Chart | Basic R Guide for NSC Statistics [Internet]. [cited 2026 Jan 29]. Available from: https://bookdown.org/dli/rguide/pie-chart.html
12. 	Pie chart with percentages in ggplot2 [Internet]. R CHARTS | A collection of charts and graphs made with the R programming language. 2021 [cited 2026 Jan 29]. Available from: https://r-charts.com/part-whole/pie-chart-percentages-ggplot2/
13. 	OR R. Results from rfe function (caret) to compute average metrics - R [Internet]. Cross Validated. 2016 [cited 2026 Feb 1]. Available from: https://stats.stackexchange.com/q/214387
14. 	Compare The Performance of Machine Learning Algorithms in R - MachineLearningMastery.com [Internet]. [cited 2026 Feb 1]. Available from: https://machinelearningmastery.com/compare-the-performance-of-machine-learning-algorithms-in-r/
15. 	r - Why does plot gives wrong score of fits using MLeval package? - Stack Overflow [Internet]. [cited 2026 Feb 1]. Available from: https://stackoverflow.com/questions/67505412/why-does-plot-gives-wrong-score-of-fits-using-mleval-package
16. 	Pie [Internet]. [cited 2026 Feb 1]. Available from: https://plotly.com/r/pie-charts/
17. 	Heatmaps [Internet]. [cited 2026 Feb 1]. Available from: https://plotly.com/r/heatmaps/
18. 	John CR. MLeval: Machine Learning Model Evaluation [Internet]. 2020 [cited 2026 Feb 2]. Available from: https://cran.r-project.org/web/packages/MLeval/index.html

